Community Embeddings for Friend Suggestions

### 2020-10-11 conference_101719.tex
paste in alg stuff from MT

\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithm}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage[style=numeric, backend=bibtex, language=english, maxbibnames=99]{biblatex}

@@

>>>>

\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage[style=numeric, backend=bibtex, language=english, maxbibnames=99]{biblatex}

@@

\begin{algorithm}[H]
    \caption{Top-$N$ Social Recommendations based on Node Embeddings}
    \label{alg:friend_sug}
    \begin{algorithmic}[1]
        \Require{graph $G = (V, E)$, maximum number of communities K, number of walks $\gamma$, walk length $\ell$, window size $\zeta$, representation size $D$, negative context size $m$, parameters ($\alpha$, $\beta$), number of recommendations $N$.}
        \Ensure{Top-$N$ recommendations for all nodes $R$.}
        % run ComE BGMM+VI
        \State $\Phi, \Phi', \Pi, (\Psi,\Sigma) \leftarrow ComE_{BGMM+VI}(G, K, \gamma, \ell, \zeta, D, m, \alpha, \beta)$ \Comment{Run ComE BGMM+VI.}
        % find friends
        \For{$(v, v') \in E$} \Comment{Determine sets of friends for each user.}
        \State $F_v \leftarrow F_v \cup \{v'\}$
        \State $F_{v'} \leftarrow F_{v'} \cup \{v\}$
        \EndFor
        % for each node
        \For{$v \in V$}
        % generate ascending recommendation list for each node
        \State $R_v \leftarrow SortedDict(size=N)$ \Comment{Init recommendations for node $v$.}
        \For{$v' \in V \land v' \notin F_v \land v' \neq v$} \Comment{For all non-friends of user $v$.}
        \State $d \leftarrow \|\phi_v - \phi_{v'}\|$ \Comment{Distance of $v$ and $v'$ embeddings.}
        \State $R_v[d] \leftarrow v'$ \Comment{Add $v'$ to $v$'s recommendations at distance $d$.}
        \EndFor
        % get Top-N recommendations for each node
        \State $R_v  \leftarrow R_v.values()$ \Comment{Get top-$N$ recommendations for each node.}
        \EndFor
    \end{algorithmic}
\end{algorithm}

Algorithm \ref{alg:friend_sug} describes in pseudocode how recommendations are computed from node embeddings generated by ComE BGMM+VI. The terms friend suggestions and social recommendations are used interchangeably.

The referenced function in line 1 $ComE_{BGMM+VI}(G, \gamma, \ell, \zeta, D, m, \alpha, \beta)$ is the ComE BGMM+VI algorithm as described in Section \ref{sec:ComE_BGMM}. ComE BGMM+VI is based on the ComE algorithm presented by \citeauthor{ComE} in their paper \citetitle{ComE} as algorithm 1 on page 381 of the publication's in-proceeding. The runtime of ComE BGMM+VI is equivalent to the runtime of ComE in it's big-$O$ notation. ComE's time complexity is $O(\abs{V}\gamma\ell + \abs{V} + T_1 (T_2\abs{V}K + K + \abs{E} + \abs{V}\gamma\ell + \abs{V}K))$, which is linear in time complexity to the graph size: $O(\abs{V} + \abs{E})$.\cite{ComE} $T_1$ is the number of iterations of ComE's closed loop consisting of node embedding, community detection, and community embedding to run. $T_2$ is the number of iterations to run expectation-maximization or variational inference for community detection and embedding.

ComE uses a Gaussian mixture model and expectation-maximization for community embedding, therefore heavily relying on the number of communities $K$ hyperparameter for embedding quality. ComE BGMM+VI instead takes a Bayesian approach by utilizing a Bayesian Gaussian Mixture with variational inference for community embeddings. variational inference allows dropping unused communities and therefore is less dependent on the number of communities $K$ hyperparameter. In their paper \citetitle{ComE+}, \citeauthor{ComE+} present a Bayesian approach to ComE as algorithm 45 of their publication's magazine. ComE+'s time complexity is $O(\abs{V}\gamma\ell + \abs{V} + T_1 (T_2\abs{V}K + K + \abs{E} + \abs{V}\gamma\ell + \abs{V}K))$ as well, which is linear in time complexity to the graph size: $O(\abs{V} + \abs{E})$.\cite{ComE+}

The function $ComE_{BGMM+VI}(G, K, \gamma, \ell, \zeta, D, m, \alpha, \beta)$ returns the node embedding $\Phi$, context embedding $\Phi'$, community assignment $\Pi$, and community embedding $(\Psi,\Sigma)$.

Let us consider the full algorithm's time complexity now:
\begin{enumerate}
    \item Line 1: The node and community embeddings are generated with ComE BGMM+VI in linear time to the graph size: $O(\abs{V}\gamma\ell + \abs{V} + T_1 (T_2\abs{V}K + K + \abs{E} + \abs{V}\gamma\ell + \abs{V}K)) = O(\abs{V}\abs{E})$.
    \item Lines 2-5: All friends of each user are determined in $O(E)$.
    \item Lines 6 through 13 compute a dictionary of the Top-$N$ friend suggestions for each user sorted ascending by distances between the node embeddings. Each node is considered: $O(\abs{V})$.
          \begin{enumerate}
              \item Line 7: A SortedDict is initialized in $O(1)$.
              \item Lines 8-11: Iterating over all non-friends $v'$ of user $v$ is achieved in $O(\abs{V})$. The distance in embedding between user $v$ and $v'$ are calculated in line 9 in $O(1)$. In line 10, friend suggestions are added to user $v$'s sorted friends suggestions in $O(N)$. The number of friends per user $\overline{\abs{F_v}}$ is not considered, because $\overline{\abs{F_v}} \ll \abs{V}$. Lines 8-11 are therefore achievable in $O(\abs{V}N)$.
              \item Line 12: Friend suggestions are returned as lists of top-$N$ users in $O(1)$.
          \end{enumerate}
          Lines 6 through 13 are achieved in $O(\abs{V}^2N))$
\end{enumerate}

According to the analysis in the list above, the total time complexity of Algorithm \ref{alg:friend_sug} is $O(\abs{V}\gamma\ell + \abs{V} + T_1 (T_2\abs{V}K + K + \abs{E} + \abs{V}\gamma\ell + \abs{V}K) + \abs{E} + \abs{V}^2N)$. We consider $\gamma$, $\ell$, $T_1$, $T_2$, and $N$ as constant, therefore the resulting time complexity in big-O notation depends on the graph's size $(\abs{V}, \abs{E})$ and the number of communities $K$: $O(K + \abs{V}K + \abs{E} + \abs{V}^2N)$. The number of communities $K$ can also be considered as constant, in which case the time complexity is $O(\abs{E} + \abs{V}^2N)$.

\begin{algorithm}[H]
    \caption{Top-$N$ Social Recommendations based on Node and Community Embeddings}
    \label{alg:friend_sug_com}
    \begin{algorithmic}[1]
        \Require{graph $G = (V, E)$, maximum number of communities K, number of walks $\gamma$, walk length $\ell$, window size $\zeta$, representation size $D$, negative context size $m$, parameters ($\alpha$, $\beta$), number of recommendations $N$.}
        \Ensure{Top-$N$ recommendations for all nodes $R$.}
        % run ComE BGMM+VI
        \State $\Phi, \Phi', \Pi, (\Psi,\Sigma) \leftarrow ComE_{BGMM+VI}(G, K, \gamma, \ell, \zeta, D, m, \alpha, \beta)$ \Comment{Run ComE BGMM+VI.}
        %  generate a set of nodes for each community
        \For{$v \in V$} \Comment{Generate nodes sets $C_k$ per community $k$.}
        \State $k \leftarrow \pi_v$
        \State $C_k \leftarrow C_k \cup \{v\}$
        \EndFor
        % find friends
        \For{$(v, v') \in E$} \Comment{Determine sets of friends for each user.}
        \State $F_v \leftarrow F_v \cup \{v'\}$
        \State $F_{v'} \leftarrow F_{v'} \cup \{v\}$
        \EndFor
        % for each node
        \For{$v \in V$}
        % generate ascending recommendation list for each node
        \State $R_v \leftarrow SortedDict(size=N)$ \Comment{Init recommendations for node $v$.}
        \For{$v' \in C_k \land v' \notin F_v \land v' \neq v$} \Comment{For all non-friends of user $v$.}
        \State $d \leftarrow \|\phi_v - \phi_{v'}\|$ \Comment{Distance of $v$ and $v'$ embeddings.}
        \State $R_v[d] \leftarrow v'$ \Comment{Add $v'$ to $v$'s recommendations at distance $d$.}
        \EndFor
        % get Top-N recommendations for each node
        \State $R_v  \leftarrow R_v.values()$ \Comment{Get top-$N$ recommendations for each node.}
        \EndFor
    \end{algorithmic}
\end{algorithm}

Algorithm \ref{alg:friend_sug_com} describes in pseudocode how recommendations are computed from ComE BGMM+VI node and community embeddings and community assignments. The advantage of the second variant of the algorithm, which also considers communities, is that an improvement in time complexity can be obtained, by considering only users in the same community for friend suggestions.

Line 1, lines 6-9, and lines 10-17, except for a small edit in line 12, are the same as in Algorithm \ref{alg:friend_sug}. Differences to Algorithm \ref{alg:friend_sug}:

\begin{enumerate}
    \item Lines 2-5: All users are filtered into sets $C_k$, one for each community $k$ in $O(\abs{V})$.
    \item Line 12: Instead of considering the set of all non-friends users of size $\abs{V}$, only users in the same community $C_k$ of size $\frac{\abs{V}}{K}$ are considered. Lines 10 through 17 are achievable in $O(\frac{\abs{V}^2N}{K})$.
\end{enumerate}

According to the analysis from Algorithm \ref{alg:friend_sug} and the differences to Algorithm \ref{alg:friend_sug_com} above, the total time complexity of Algorithm \ref{alg:friend_sug_com} is updated to $O(\abs{V}\gamma\ell + \abs{V} + T_1 (T_2\abs{V}K + K + \abs{E} + \abs{V}\gamma\ell + \abs{V}K) + \abs{V} + \abs{E} + \frac{\abs{V}^2N}{K})$. Again, we consider $\gamma$, $\ell$, $T_1$, $T_2$, and $N$ as constant, therefore the resulting time complexity in big-O notation depends on the graph's size $(\abs{V}, \abs{E})$ and the number of communities $K$: $O(K + \abs{V}K + \abs{E} + \frac{\abs{V}^2N}{K})$. The number of communities $K$ can also be considered as constant, in which case the time complexity is $O(\abs{E} + \abs{V}^2N)$.

The time complexity of Algorithm \ref{alg:friend_sug_com}, when not considering $K$ as a substantial factor in time complexity is equivalent to the time complexity of Algorithm \ref{alg:friend_sug}. Nevertheless, it is clear, that $K$ can substantially reduce the quadratic summand $\abs{V}^2N$ through division by only considering users of the same community for friend suggestions.
